Hyper-linked Communications: WebRTC enabled asynchronous collaboration

### 2016-05-13 tese2/07_related_work_nat.tex
until 1h test case

The need to build a global communications network in an age when almost nobody had access to that technology and the unpredictability of the number of future users, lead to some protocols not being suitable for the explosive growth and proliferation of users that followed. \ac{IPv4} limits the number of public addresses in such a way that today they are scarce \cite{ipv4}. Today we live in the \ac{IoT} era.
  Users and companies may have multiple devices that require Internet connectivity, such as personal computers, smart phones, tablets, traffic sensors, public surveilance systems, health monitoring tools and so on. All those devices combined cannot be publicly addressed by \ac{IPv4}.
%RP vigillance -> surveilance?
%RP usa acrónimo IoT para Internet of Things

@@

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.7\textwidth]{figures/nat.pdf}
	\caption{Network Address Translation}
	\label{fig:nat}
\end{figure}
%RP na figura devias indicar que uns tem ips publicos e outros privados. Alternativamente, colocavas metade da nat box dentro da núvem para indicar que esta pertence à Internet e os pc não.

@@

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.45\textwidth]{figures/multinat.pdf}
	\caption{Multiple level NATs}
	\label{fig:multinat}
\end{figure}

@@

On the other hand, \ac{TURN} uses public servers to relay traffic between private endpoints.
It may use a \ac{P2P} network relay to find the best peer, but after that, the behavior is much like client-server. Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \emph{full cone}. \ac{ICE} is a technique that uses \ac{STUN} when direct communications are possible and \ac{TURN} while a direct communication isn't possible.
%RP explicar que para usar TURN as aplicações têm de estar cientes disso.

Most of client-server applications aren't affected by \ac{NAT} when the servers are public, but they're inadequate for real time communication between two private endpoints. Clearly \ac{TURN} requires a more expensive relaying infrastructure and, in most cases, more network usage, leading to a worse quality of service. The requirements of real-time video communication makes this kind of model unsuitable.

>>>>

The need to build a global communications network in an age when almost nobody had access to that technology and the unpredictability of the number of future users, lead to some protocols not being suitable for the explosive growth and proliferation of users that followed. \ac{IPv4} limits the number of public addresses in such a way that today they are scarce \cite{ipv4}. Today we live in the \ac{IoT} era.
  Users and companies may have multiple devices that require Internet connectivity, such as personal computers, smartphones, tablets, traffic sensors, public surveilance systems, health monitoring tools and so on. All those devices combined cannot be publicly addressed by \ac{IPv4}.
%RP vigillance -> surveilance?
%RP usa acrónimo IoT para Internet of Things

@@

\begin{figure}
	\centering
	\includegraphics[width=0.7\textwidth]{figures/nat.pdf}
	\caption{Network Address Translation}
	\label{fig:nat}
\end{figure}
%RP na figura devias indicar que uns tem ips publicos e outros privados. Alternativamente, colocavas metade da nat box dentro da núvem para indicar que esta pertence à Internet e os pc não.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{figures/multinat.pdf}
	\caption{Multiple level NATs}
	\label{fig:multinat}
\end{figure}

@@

On the other hand, \ac{TURN} uses public servers to relay traffic between private endpoints.
It may use a \ac{P2P} network relay to find the best peer, but after that, the behavior is much like client-server. Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \emph{full cone}. \ac{ICE} is a technique that uses \ac{STUN} when direct communications are possible and \ac{TURN} while a direct communication is not possible.
%RP explicar que para usar TURN as aplicações têm de estar cientes disso.

Most of client-server applications are not affected by \ac{NAT} when the servers are public, but they're inadequate for real time communication between two private endpoints. Clearly \ac{TURN} requires a more expensive relaying infrastructure and, in most cases, more network usage, leading to a worse quality of service. The requirements of real-time video communication makes this kind of model unsuitable.

### 2016-05-13 tese2/08_related_work_rtc.tex
[ADD] template2
[ADD] amazing explanation of network test results
[ADD] references to tables
[FIX] reference to stroph.js
[RFT] removed some blue text and some old RP comments
until 1h test case

\subsection{Real time communications}\label{rtc}

\ac{WebRTC} is an open source technology that defines a collection of standard protocols and JavaScript \ac{API}s for web browser based real time communications without installing any additional application or plug-in.

@@

%RP indicar que agora também já temos webrtc fora do browser devido ao seu sucesso e ao facto de reunir tecnologia state of the art?

\newcolumntype{Y}{>{\centering\arraybackslash}X}
%\renewcommand{\arraystretch}{2}

\begin{table}[]
\centering
\label{my-label}
    \begin{tabularx}{\textwidth}{*8{Y}@{}m{0pt}@{}} 
    	\cline{5-8}
		&&&\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{~~~~~~MediaStream~~~~~~} & \multicolumn{2}{c|}{~~~~~~DataChannel~~~~~~} 										&\\[12pt]
		\hline 
		\multicolumn{1}{|c|}{XHR} & \multicolumn{1}{c|}{SSE} & \multicolumn{2}{c|}{WebSocket} & \multicolumn{2}{c|}{SRTP} & \multicolumn{2}{c|}{SCTP} 				&\\[12pt]
		\cline{1-3}\cline{6-8} 
		\multicolumn{3}{|c|}{HTTP 1.X/2.0} & \multicolumn{1}{c|}{~~~~~~~~~~~~~~~~~~} & \multicolumn{1}{c}{~~~~~~~~~~~~~~~~} & \multicolumn{3}{|c|}{Session (DTLS)} 	&\\[12pt]
		\hline
		\multicolumn{4}{|c|}{Session (TLS) - optional} & \multicolumn{4}{c|}{ICE, STUN, TURN} 																		&\\[12pt]
		\hline
		\multicolumn{4}{|c|}{Transport (TCP)} & \multicolumn{4}{c|}{Transport (UDP)} 																				&\\[12pt]
		\hline
		\multicolumn{8}{|c|}{Network (IP)}  																														&\\[12pt]
		\hline
	\end{tabularx}
	\caption{WebRTC protocol Stack}

@@

  \item \textbf{PeerConnection} acquires connection data and negotiates with peers.
%RP establishes and negociates a connection with a peer for transmitting real-time video or audio?
  \item \textbf{DataChannel} provides a channel for exchanging arbitrary data with other peers.
\end{itemize}

@@

\ac{TLS} doesn't support independent packet decryption\cite{rfc6347}, for that it requires a reliable transport channel, typically \ac{TCP}. The decryption of a packet depends on the previous packet, which for unreliable transport protocols like \ac{UDP} may represent a problem, either due to packet loss or different reception order.

\ac{DTLS} is similar to \ac{TLS}, but is used on top of \ac{UDP}.
The main difference is the inclusion of a sequence number per packet that is used for packet re-ordering on reception and protects from duplicated packets. If a packet sequence number is less than the expected sequence number the packet is discarded. If a packet sequence number is greater than the expected sequence number the packet may be enqueued or discarded. By knowing the sequence of messages that are sent and received in \ac{DTLS}, timers are used for packet retransmission avoiding acknowledgment messages.
%RP TLS?
%\def\tabularxcolumn#1{m{#1}}
\begin{table}[]
\centering
    \begin{tabularx}{\textwidth}{X|Y|Y|Y|@{}m{0pt}@{}}
\cline{2-4}
                                                     & \textbf{TCP}  & \textbf{UDP}     & \textbf{SCTP}    &\\[12pt]\hline
\multicolumn{1}{|c|}{\textbf{Reliability}}           & reliable      & unreliable       & configurable     &\\[12pt]\hline
\multicolumn{1}{|c|}{\textbf{Delivery}}              & ordered       & unordered        & configurable     &\\[12pt]\hline
\multicolumn{1}{|c|}{\textbf{Transmission}}          & byte-oriented & message-oriented & message-oriented &\\[12pt]\hline
\multicolumn{1}{|c|}{\textbf{Flow control}}          & yes           & no               & yes              &\\[12pt]\hline
\multicolumn{1}{|c|}{\textbf{Congestion control}} 	 & yes           & no               & yes              &\\[12pt]\hline
\end{tabularx}
	\caption{Overview of transport protocols}
\end{table}

\ac{WebRTC}'s \emph{DataChannel} is built on top of \ac{SCTP}, which is encapsulated by \ac{DTLS}. \ac{DTLS} encapsulation provides confidentiality, authentication and integrity to the transfered data. A \emph{Data Channel} has one incoming stream and one outgoing stream, providing bidirectional communication. Each data channel direction can be configured for reliable or unreliable transmission, the same can be done for order delivery and priority. which can also be defined for improving the quality of service of a particular stream over the others.

\ac{WebRTC}'s \emph{MediaStream} is built on top of \ac{SRTP}, which requires an external mechanism for key exchange. \ac{DTLS} keys are negotiated on handshake in order to achieve a secure connection. The new keys derived from \ac{DTLS} handshake are seized for \ac{SRTP} encryption, the remaining \ac{SRTP} communications are done through \ac{UDP} without using \ac{DTLS}.
%RP quais remaining?  A última frase é confusa

\ac{WebRTC} aims to provide a standard platform for real-time audio and video on the Web. It arrives at a time when several proprietary products are well established.
\emph{Skype}\footnote{\url{http://www.skype.com/}(accessed June 1, 2015).} is an application that allows video, voice, instant messaging and multi-party communication over proprietary protocols, its main strength are the amount of users that are using it nowadays and the ability to perform voice calls to the \ac{PSTN}. But compared to \emph{Skype}, \ac{WebRTC} applications don't need to be pre-installed.

\emph{Google Hangouts}\footnote{\url{http://plus.google.com/hangouts}(accessed June 1, 2015).} is another popular video multi-party conference web application. 
In the past, in order to use \emph{hangouts} on a web browser a plug-in had to be installed, nowadays hangouts is using \ac{WebRTC}. \emph{Google Hangouts} supports viewing videos on \emph{youtube} synchronously, drawing collaboratively, creating music and playing multi-player games. These applications are implemented with \emph{Adobe Flash}.

>>>>

\section{Real time communications}\label{rtc}

\ac{WebRTC} is an open source technology that defines a collection of standard protocols and \emph{JavaScript} \ac{API}s for web browser based real time communications without installing any additional application or plug-in. Table \ref{table:webrtcstack} shows the protocols that \ac{WebRTC} rely on.

@@

%\renewcommand{\arraystretch}{2}

\begin{table}
\centering
\caption{WebRTC protocol Stack}
\label{table:webrtcstack}
    \begin{tabularx}{\textwidth}{*8{Y}@{}m{0pt}@{}} 
    	\cline{5-8}
		&&&\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{~~~~~~~~~~MediaStream~~~~~~~~~~} & \multicolumn{2}{c|}{~~~~~~~~~~DataChannel~~~~~~~~~~} 								&\\[12pt]
		\hline 
		\multicolumn{1}{|c|}{XHR} & \multicolumn{1}{c|}{SSE} & \multicolumn{2}{c|}{WebSocket} & \multicolumn{2}{c|}{SRTP} & \multicolumn{2}{c|}{SCTP} 				&\\[12pt]
		\cline{1-3}\cline{6-8} 
		\multicolumn{3}{|c|}{HTTP 1.X/2.0} & \multicolumn{1}{c|}{~~~~~~~~~~~~~~~~~~} & \multicolumn{1}{c}{~~~~~~~~~~~~~~~~} & \multicolumn{3}{|c|}{Session (DTLS)} 	&\\[12pt]
		\hline
		\multicolumn{4}{|c|}{Session (TLS) - optional} & \multicolumn{4}{c|}{ICE, STUN, TURN} 																		&\\[12pt]
		\hline
		\multicolumn{4}{|c|}{Transport (TCP)} & \multicolumn{4}{c|}{Transport (UDP)} 																				&\\[12pt]
		\hline
		\multicolumn{8}{|c|}{Network (IP)}  																														&\\[12pt]
		\hline
	\end{tabularx}

@@

  \item \textbf{PeerConnection} acquires connection data and negotiates with peers.

  \item \textbf{DataChannel} provides a channel for exchanging arbitrary data with other peers.
\end{itemize}

@@

\ac{TLS} does not support independent packet decryption\cite{rfc6347}, for that it requires a reliable transport channel, typically \ac{TCP}. The decryption of a packet depends on the previous packet, which for unreliable transport protocols like \ac{UDP} may represent a problem, either due to packet loss or different reception order.

\ac{DTLS} is similar to \ac{TLS}, but is used on top of \ac{UDP}.
The main difference is the inclusion of a sequence number per packet that is used for packet re-ordering on reception and protects from duplicated packets. If a packet sequence number is less than the expected sequence number the packet is discarded. If a packet sequence number is greater than the expected sequence number the packet may be enqueued or discarded. By knowing the sequence of messages that are sent and received in \ac{DTLS}, timers are used for packet retransmission avoiding acknowledgment messages.

Table \ref{table:transport_protocols} shows an overview of the transport protocols.
\ac{WebRTC}'s \emph{DataChannel} is built on top of \ac{SCTP}, which is encapsulated by \ac{DTLS}. \ac{DTLS} encapsulation provides confidentiality, authentication and integrity to the transfered data. A \emph{Data Channel} has one incoming stream and one outgoing stream, providing bidirectional communication. Each data channel direction can be configured for reliable or unreliable transmission, the same can be done for order delivery and priority. which can also be defined for improving the quality of service of a particular stream over the others.

%RP TLS?
%\def\tabularxcolumn#1{m{#1}}
\begin{table}
\centering
	\caption{Overview of transport protocols}
	\label{table:transport_protocols}
    \begin{tabularx}{\textwidth}{X|Y|Y|Y|@{}m{0pt}@{}}
\cline{2-4}
                                                     & \textbf{TCP}  & \textbf{UDP}     & \textbf{SCTP}    &\\[12pt]\hline
\multicolumn{1}{|c|}{\textbf{Reliability}}           & reliable      & unreliable       & configurable     &\\[12pt]\hline
\multicolumn{1}{|c|}{\textbf{Delivery}}              & ordered       & unordered        & configurable     &\\[12pt]\hline
\multicolumn{1}{|c|}{\textbf{Transmission}}          & byte-oriented & message-oriented & message-oriented &\\[12pt]\hline
\multicolumn{1}{|c|}{\textbf{Flow control}}          & yes           & no               & yes              &\\[12pt]\hline
\multicolumn{1}{|c|}{\textbf{Congestion control}} 	 & yes           & no               & yes              &\\[12pt]\hline
\end{tabularx}
\end{table}

\ac{WebRTC}'s \emph{MediaStream} is built on top of \ac{SRTP}, which requires an external mechanism for key exchange. \ac{DTLS} keys are negotiated on handshake in order to achieve a secure connection. The new keys derived from \ac{DTLS} handshake are seized for \ac{SRTP} encryption, the remaining \ac{SRTP} communications are done through \ac{UDP} without using \ac{DTLS}.

\ac{WebRTC} aims to provide a standard platform for real-time audio and video on the Web. It arrives at a time when several proprietary products are well established.
\emph{Skype}\footnote{\url{http://www.skype.com/}(accessed June 1, 2015).} is an application that allows video, voice, instant messaging and multi-party communication over proprietary protocols, its main strength are the amount of users that are using it nowadays and the ability to perform voice calls to the \ac{PSTN}. But compared to \emph{Skype}, \ac{WebRTC} applications do not need to be pre-installed.

\emph{Google Hangouts}\footnote{\url{http://plus.google.com/hangouts}(accessed June 1, 2015).} is another popular video multi-party conference web application. 
In the past, in order to use \emph{hangouts} on a web browser a plug-in had to be installed. Nowadays hangouts is using \ac{WebRTC}. \emph{Google Hangouts} supports viewing videos on \emph{youtube} synchronously, drawing collaboratively, creating music and playing multi-player games. These applications are implemented with \emph{Adobe Flash}.

### 2016-05-13 tese2/09_related_work_signaling.tex
[RFT] removed some blue text and some old RP comments

  \ac{WebRTC} uses \ac{SDP} \cite{rfc4566} to define peer connection properties such as types of supported media, codecs, protocols used and network information. An \ac{SDP} offer describes to other peers the expected type of communication and its details, such as used transport protocols, codecs, security and other.

  One of \ac{WebRTC} signaling's requisites is bi-directional communication.
%RP acrescentei o webrtc pois isto não é um requisito geral de signaling. era o que querias dizer?
%HR sim é isso
  \ac{HTTP} uses a request-response paradigm, where a request is sent by the client, followed by a server response.
  %in other words, it follows an unidirectional communication.
%RP de certeza que queres dizer que o http é unidirecional? 
%HR era mais o facto de o servidor não poder comunicar com o cliente sem que o cliente comunique primeiro (tem que haver uma iniciativa por parte do cliente)
  Sometimes it is required that some information be obtained in real time, but we saw, some \ac{NAT}'s do not support callbacks from servers, preventing them from notifying clients as soon as an event occures.
One technique to overcome this problem is polling.
  %RP num documento formal não se usam contrações como it's.
  %RP o teu ingles é parco em artigos. Faltam muitos the, of, it, etc.

Polling consists on sending periodic messages to which the server responds immediately with empty content or fresh information. Text and presence messages are unpredictable, if the time between periodic requests is short, most of the time the server will return empty results wasting network bandwidth and energy. On the other hand, if the time between periodic requests is large, newer messages may arrive too late.
%RP  Real time communications are unpredictable. Normalmente falamos de audio e video, que são previsíveis. Estás a referir-te às comunicações em si ou ao signaling e estabelecimento?

@@

The \emph{WebSocket} protocol allows bi-directional communications over a full-duplex socket channel \cite{rfc6455}, by other words it supports sending and receiving data simultaneously.
%RP coloca as citações no final da frase. Pode haver excepções onde a citação apenas cobre parte da frase ou precisas de colocar várias a dizer coisas diferentes.
 \emph{WebSocket} handshake phase specifies a \ac{HTTP} header in order to upgrade to \emph{WebSocket} type of communication, but the remainder messages are exchanged without \ac{HTTP} headers, which leads to much smaller messages and better network usage. WebSockets may not be available on every web browser, frameworks like \emph{socket.io} and \emph{SockJS}\footnote{\url{http://github.com/sockjs}(accessed June 1, 2015).} fall back to using \ac{HTTP} long polling when there is no support for \emph{WebSockets}.

  \ac{BOSH}\cite{xep0124} is a technique based on long polling that uses two socket connections and allows sending client messages to the server while a previous request is held.
  The \ac{BOSH} specification assumes that a connection manager is implemented to handle \ac{HTTP} connections. This connection manager is basically a translator from \ac{HTTP} to raw message so that the server may be implemented as if this communication is performed over \ac{TCP}.
  When the connection manager holds for a response for too long, it responds with an empty body, this technique prevents an \ac{HTTP} session from expiring when the client is waiting for a response, thus expanding the session time. Expiring sessions can be expensive due to the overhead of establishing new connections, which is even worse when \ac{HTTP} is used over \ac{SSL}.

@@

  \ac{SIP} \cite{rfc3261} is a protocol used for negotiation, creation, modification and finalization of communication sessions between users. \ac{SIP} follows a client/server architecture with \ac{HTTP} like messages and it can be used as a signaling protocol. The advantage of \ac{SIP} is the ability to make video and voice calls between the telephone network and applications over \ac{IP} networks.

  The working group \ac{SIMPLE}\footnote{\url{https://datatracker.ietf.org/wg/simple/documents/}(accessed June 1, 2015).} proposed the creation of \ac{SIP} extensions, namely presence information \cite{rfc5263} and instant messaging \cite{rfc3428}.

@@

  \ac{XMPP} was initially developed for instant messaging and presence (Jabber\footnote{\url{http://jabber.org/}(accessed June 1, 2015).}). It is nowadays an open technology for standardized, decentralized, secure and extensible real-time communications. 
  \ac{XMPP} messages are \ac{XML} based, which is attractive for applications that need structured messages and rich hypermedia. Another advantage of \ac{XMPP} is the addition of extensions, for example \emph{XEP-0096}\cite{xep0096}, which adds file transfer capabilities between two entities, and \emph{XEP-0045}\cite{xep0045}, which enables multi-user chat.
  \ac{XMPP}'s bi-directional communication over \ac{HTTP} is achieved through \ac{BOSH} \cite{xep0206}.
  This kind of communication is also possible through WebSockets \cite{rfc7395}.
  Today, multiple \ac{XMPP} server implementations exists, such as: \emph{ejabberd}\footnote{\url{http://jabberd.im/}(accessed June 1, 2015).}, \emph{Metronome}\footnote{\url{http://lightwitch.org/metronome}(accessed June 1, 2015).}, \emph{Openfire}\footnote{\url{http://igniterealtime.org/projects/openfire/}(accessed June 1, 2015).} and \emph{Prosody}\footnote{\url{http://prosody.im/}(accessed June 1, 2015).}. \emph{Ejabberd} is the server that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}(accessed June 1, 2015).}.
  %RP fazes muitos paragrafos. Um paragrafo por tecnologia é normalmente suficiente se não forem muito grandes.

{\color{blue}

@@

    The user profile is represented in terms of Data Forms\cite{xep0004} protocol and managed (registration and update) through \textit{Info/Query} (IQ) \cite{rfc3920} requests to the \ac{XMPP} server, which responds with an IQ response containing information to retrieve the registration fields needed to fill its profile (eg username, password, telephone). 
%RP frase demasiado longa. Torna-se confusa.

@@

    We know that an \ac{XMPP} based application could simplify our work just by using functionalities that \ac{XMPP} already implement, but  we don't need all of them. In fact, we just need a subset of \ac{XMPP} features, namely a simple way to register users, access and edit user's information, create chat rooms, send messages and access to presence information.
}

>>>>

  \ac{WebRTC} uses \ac{SDP}~\cite{rfc4566} to define peer connection properties such as types of supported media, \emph{codecs}, protocols used and network information. An \ac{SDP} offer describes to other peers the expected type of communication and its details, such as used transport protocols, codecs, security and other.

  One of \ac{WebRTC} signaling's requisites is bi-directional communication.
%RP acrescentei o webrtc pois isto não é um requisito geral de signaling. era o que querias dizer?
%HR sim é isso
  \ac{HTTP} uses a request-response paradigm, where a request is sent by the client, followed by a server response.
  %in other words, it follows an unidirectional communication.
%RP de certeza que queres dizer que o http é unidirecional? 
%HR era mais o facto de o servidor não poder comunicar com o cliente sem que o cliente comunique primeiro (tem que haver uma iniciativa por parte do cliente)
  Sometimes it is required that some information be obtained in real time, but we saw, some \ac{NAT}'s do not support callbacks from servers, preventing them from notifying clients as soon as an event occures.
One technique to overcome this problem is polling. Polling consists on sending periodic messages to which the server responds immediately with empty content or fresh information. Text and presence messages are unpredictable, if the time between periodic requests is short, most of the time the server will return empty results wasting network bandwidth and energy. On the other hand, if the time between periodic requests is large, newer messages may arrive too late.
%RP  Real time communications are unpredictable. Normalmente falamos de audio e video, que são previsíveis. Estás a referir-te às comunicações em si ou ao signaling e estabelecimento?

@@

The \emph{WebSocket} protocol allows bi-directional communications over a full-duplex socket channel~\cite{rfc6455}, by other words it supports sending and receiving data simultaneously.
%RP coloca as citações no final da frase. Pode haver excepções onde a citação apenas cobre parte da frase ou precisas de colocar várias a dizer coisas diferentes.
 \emph{WebSocket} handshake phase specifies a \ac{HTTP} header in order to upgrade to \emph{WebSocket} type of communication, but the remainder messages are exchanged without \ac{HTTP} headers, which leads to much smaller messages and better network usage. WebSockets may not be available on every web browser, frameworks like \emph{socket.io} and \emph{SockJS}\footnote{\url{http://github.com/sockjs}(accessed June 1, 2015).} fall back to using \ac{HTTP} long polling when there is no support for \emph{WebSockets}.

  \ac{BOSH}~\cite{xep0124} is a technique based on long polling that uses two socket connections and allows sending client messages to the server while a previous request is held.
  The \ac{BOSH} specification assumes that a connection manager is implemented to handle \ac{HTTP} connections. This connection manager is basically a translator from \ac{HTTP} to raw message so that the server may be implemented as if this communication is performed over \ac{TCP}.
  When the connection manager holds for a response for too long, it responds with an empty body, this technique prevents an \ac{HTTP} session from expiring when the client is waiting for a response, thus expanding the session time. Expiring sessions can be expensive due to the overhead of establishing new connections, which is even worse when \ac{HTTP} is used over \ac{SSL}.

@@

  \ac{SIP}~\cite{rfc3261} is a protocol used for negotiation, creation, modification and finalization of communication sessions between users. \ac{SIP} follows a client/server architecture with \ac{HTTP} like messages and it can be used as a signaling protocol. The advantage of \ac{SIP} is the ability to make video and voice calls between the telephone network and applications over \ac{IP} networks.

  The working group \ac{SIMPLE}\footnote{\url{https://datatracker.ietf.org/wg/simple/documents/}(accessed June 1, 2015).} proposed the creation of \ac{SIP} extensions, namely presence information~\cite{rfc5263} and instant messaging~\cite{rfc3428}.

@@

  \ac{XMPP} was initially developed for instant messaging and presence (Jabber\footnote{\url{http://jabber.org/}(accessed June 1, 2015).}). It is nowadays an open technology for standardized, decentralized, secure and extensible real-time communications. 
  \ac{XMPP} messages are \ac{XML} based, which is attractive for applications that need structured messages and rich hypermedia. Another advantage of \ac{XMPP} is the addition of extensions, for example \emph{XEP-0096}~\cite{xep0096}, which adds file transfer capabilities between two entities, and \emph{XEP-0045}~\cite{xep0045}, which enables multi-user chat.
  \ac{XMPP}'s bi-directional communication over \ac{HTTP} is achieved through \ac{BOSH}~\cite{xep0206}.
  This kind of communication is also possible through WebSockets~\cite{rfc7395}.
  Today, multiple \ac{XMPP} server implementations exists, such as: \emph{ejabberd}\footnote{\url{http://jabberd.im/}(accessed June 1, 2015).}, \emph{Metronome}\footnote{\url{http://lightwitch.org/metronome}(accessed June 1, 2015).}, \emph{Openfire}\footnote{\url{http://igniterealtime.org/projects/openfire/}(accessed June 1, 2015).} and \emph{Prosody}\footnote{\url{http://prosody.im/}(accessed June 1, 2015).}. \emph{Ejabberd} is the server that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}(accessed June 1, 2015).}.
  %RP fazes muitos paragrafos. Um paragrafo por tecnologia é normalmente suficiente se não forem muito grandes.

@@

    The user profile is represented in terms of Data Forms~\cite{xep0004} protocol and managed (registration and update) through \textit{Info/Query} (IQ)~\cite{rfc3920} requests to the \ac{XMPP} server, which responds with an IQ response containing information to retrieve the registration fields needed to fill its profile (eg username, password, telephone). 
%RP frase demasiado longa. Torna-se confusa.

@@

    We know that an \ac{XMPP} based application could simplify our work just by using functionalities that \ac{XMPP} already implement, but  we do not need all of them. In fact, we just need a subset of \ac{XMPP} features, namely a simple way to register users, access and edit user's information, create chat rooms, send messages and access to presence information.

### 2016-05-13 tese2/10_related_work_hypermedia.tex

  Hyper-video is a kind of video that contains links to any kind of hypermedia, including links to skip part of it. An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific time in a video or audio track. \emph{HyperCafe} \cite{hypercafe} was an experimental project to expose hyper-video concepts that consisted of an interactive film that enabled switching between different conversations taking place inside a cafe.

  Detail-on-demand is a subset of hyper-video that allow us to obtain additional information about something that appears along the video, like obtaining information about a painting that appears in a particular segment. \emph{Hyper-Hitchcock}\cite{hitchcock} is an editor and player of detail-on-demand video.

@@

  \emph{HyVAL}\cite{hyval} is an \ac{XML} based language that was proposed for modeling composition, synchronization and interaction of hypermedia. HyVAL defines defines video structure, internal video and external media objects. 
  \emph{HyVAL} uses a primary video stream, around which all other elements are organized and synchronized.
  \emph{HyVAL}'s video structure object defines a structure derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically. 
  This approach is quite restrictive if we want to apply hyper-video concepts to videos that don't follow this structure. External media objects are linked by primary video, those objects can represent other videos, images, text, animation and sound.
  %RP não percebo bem a última frase. O que é primary video?

  \ac{SMIL}\cite{smil} was introduced to describe temporal behavior of multimedia content, in particular, it could be used to overlay subtitles on films. With \ac{SMIL} it is possible to synchronize multiple videos, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyper-links, among multiple other features.
  %RP synchronize multiple sections of video? vários videos apresentados em simultâneo?

@@

The \ac{DOM} is a standard \ac{API} that allows easy management of documents that are organized in a tree structure, by providing \ac{CRUD} operations over its elements and their attributes. \ac{DOM} makes it easy to inter-operate between imperative and declarative programming languages\cite{dom}.
%RP não arranjas uma referência para o dom?

@@

  SmillingWeb \cite{smillingweb} attempts to implement a cross platform multimedia player designed for \ac{SMIL} 3.0 presentations with \emph{JavaScript} and \emph{jQuery} which, unlike \cite{ambulant}, doesn't require a plug-in to be installed and shouldn't have incompatibility issues. 
  SmillingWeb already takes advantage of \ac{HTML}5 and \ac{CSS}3.
  It takes into account unsupported web browsers through the use of \emph{Modernizr}\footnote{\url{http://modernizr.com/}(accessed June 2, 2015).}, a simple \emph{JavaScript} library that may require plug-ins if new features aren't supported.  
  But SmillingWeb just implements a subset of \ac{SMIL} 3.0 and their scheduler engine loads the \ac{SMIL} file only once, which could raise problems when dealing with \ac{SMIL} changes due to real time communications.
  Another problem with SmillingWeb is pre-loading and playing elements at the correct interval of time, which is not always possible due to high latency networks leading to  pauses during playback.

@@

  \emph{JavaScript} is an imperative object-oriented language based on \emph{ECMAScript}. It is used mainly on client-side and executed by a web browser. \emph{JavaScript} has its own implementation of \ac{DOM} and one of its advantages is the ability to download and execute code on the fly without the need of pre-installed plug-ins.

@@

  Although \ac{HTML}5, \emph{JavaScript}, \ac{CSS} and \ac{WebRTC} implement some of \emph{flash}'s features, it doesn't mean that \emph{flash} will be replaced soon.
  Instead, both technologies can be used to develop rich \emph{Internet} applications.
  It is also important to note that \ac{HTML}5 is better supported in mobile devices than \emph{Adobe Flash}.

>>>>

  Hyper-video is a kind of video that contains links to any kind of hypermedia, including links to skip part of it. An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific time in a video or audio track. \emph{HyperCafe}~\cite{hypercafe} was an experimental project to expose hyper-video concepts that consisted of an interactive film that enabled switching between different conversations taking place inside a cafe.

  Detail-on-demand is a subset of hyper-video that allow us to obtain additional information about something that appears along the video, like obtaining information about a painting that appears in a particular segment. \emph{Hyper-Hitchcock}~\cite{hitchcock} is an editor and player of detail-on-demand video.

@@

  \emph{HyVAL}~\cite{hyval} is an \ac{XML} based language that was proposed for modeling composition, synchronization and interaction of hypermedia. HyVAL defines defines video structure, internal video and external media objects. 
  \emph{HyVAL} uses a primary video stream, around which all other elements are organized and synchronized.
  \emph{HyVAL}'s video structure object defines a structure derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically. 
  This approach is quite restrictive if we want to apply hyper-video concepts to videos that do not follow this structure. External media objects are linked by primary video, those objects can represent other videos, images, text, animation and sound.
  %RP não percebo bem a última frase. O que é primary video?

  \ac{SMIL}~\cite{smil} was introduced to describe temporal behavior of multimedia content, in particular, it could be used to overlay subtitles on films. With \ac{SMIL} it is possible to synchronize multiple videos, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyper-links, among multiple other features.
  %RP synchronize multiple sections of video? vários videos apresentados em simultâneo?

@@

The \ac{DOM} is a standard \ac{API} that allows easy management of documents that are organized in a tree structure, by providing \ac{CRUD} operations over its elements and their attributes. \ac{DOM} makes it easy to inter-operate between imperative and declarative programming languages~\cite{dom}.
%RP não arranjas uma referência para o dom?

@@

  SmillingWeb \cite{smillingweb} attempts to implement a cross platform multimedia player designed for \ac{SMIL} 3.0 presentations with \emph{JavaScript} and \emph{jQuery} which, unlike \cite{ambulant}, does not require a plug-in to be installed and should not have incompatibility issues. 
  SmillingWeb already takes advantage of \ac{HTML}5 and \ac{CSS}3.
  It takes into account unsupported web browsers through the use of \emph{Modernizr}\footnote{\url{http://modernizr.com/}(accessed June 2, 2015).}, a simple \emph{JavaScript} library that may require plug-ins if new features are not supported.  
  But SmillingWeb just implements a subset of \ac{SMIL} 3.0 and their scheduler engine loads the \ac{SMIL} file only once, which could raise problems when dealing with \ac{SMIL} changes due to real time communications.
  Another problem with SmillingWeb is pre-loading and playing elements at the correct interval of time, which is not always possible due to high latency networks leading to  pauses during playback.

@@

  \emph{JavaScript} is an imperative object-oriented language based on \emph{ECMAScript}. It is used mainly on client-side and executed by a web browser. \emph{JavaScript} has its own implementation of \ac{DOM} and one of its advantages is the ability to download and execute code on-the-fly without the need of pre-installed plug-ins.

@@

  Although \ac{HTML}5, \emph{JavaScript}, \ac{CSS} and \ac{WebRTC} implement some of \emph{flash}'s features, it does not mean that \emph{flash} will be replaced soon.
  Instead, both technologies can be used to develop rich \emph{Internet} applications.
  It is also important to note that \ac{HTML}5 is better supported in mobile devices than \emph{Adobe Flash}.
